### 2025-05-17 Manuscript/Manuscript.tex
Postulates

Given a quantum state known to be prepared from a finite ensemble, \gls{qsd} seeks to identify which specific state from the ensemble corresponds to the given system. In contrast, \gls{qse}~\cite{PhysicsExclusionSource} adopts the opposite perspective: it aims to determine which states from the ensemble do not correspond to the prepared state. While \gls{qsd} has been deeply studied in recent years\cite{DiscriminationArticle}, with significant advances since its inception~\cite{helstromBook}, \gls{qse} offers a complementary framework with distinct advantages.

@@

The exclusion task can be carried out under two main protocols: \gls{me} and \gls{ze}\footnote{Also known as \emph{unambiguous exclusion}.}. In the Minimum Error scenario, the goal is to minimize the probability of mistakenly excluding the actual prepared state. In contrast, the Zero Error approach seeks to exclude a state with absolute certainty, even if that means sometimes the procedure yields an inconclusive result.

@@

@@

\begin{align}\label{equationPerfectExclusionCondition}
	\lambda_{|G|} \leq \left( \sum_{i=1}^{|G|} \sqrt{\lambda_i} \right)^2.
\end{align}

@@

In Figure (\ref{FigureQSEMEGenericEigenValues3D}) we present a 3D plot for the comparison of the success probability for exclusion for ensembles of 3 states between generic and group-generated ensembles. The $x$ and $y$ axis representing de eigenvalues of the Gram matrix are analogous to previos figures. Nevertheless the $z$ axis shows the success probability for exclusion insted of the previos heatmaps. Here we can observe how the randomly generated states raise the probabilities over the group-generated values. The generation of unitary matrices that fill $U(n)$ ensures the Gram matrices we are generating comprehend the entire set of possible Gram matrices homogeniously.

@@

I am grateful to Dr. Ramón Muñoz Tapia for allowing me to be one of his pupils and for presenting this project. I am also deeply grateful to Dr. Santiago Llorens Fernández for being the lighthouse that illuminates my path throughout this journey.

>>>>

Given a quantum state known to be prepared from a finite ensemble, \gls{qsd} seeks to identify which specific state from the ensemble corresponds to the given system. In contrast, \gls{qse}~\cite{PhysicsExclusionSource} adopts the opposite perspective, it aims to determine which states from the ensemble do not correspond to the prepared state. While \gls{qsd} has been deeply studied in recent years\cite{DiscriminationArticle}, with significant advances since its inception~\cite{helstromBook}, \gls{qse} offers a complementary framework with distinct advantages.

@@

The exclusion task can be carried out under two main protocols, \gls{me} and \gls{ze}\footnote{Also known as \emph{unambiguous exclusion}.}. In the Minimum Error scenario, the goal is to minimize the probability of mistakenly excluding the actual prepared state. In contrast, the Zero Error approach seeks to exclude a state with absolute certainty, even if that means sometimes the procedure yields an inconclusive result.

@@

\subsection{Preamble}

This project is entirely situated within the quantum realm. Therefore, we begin by contextualizing some foundational postulates, notation, and elements upon which the rest of the project relies.

\textbf{Postulate I:} A quantum state provides a complete description of a physical system. In quantum mechanics, a state can be represented by a normalized vector $\ket{\psi}$ in a Hilbert space $\mathcal{H}$. Hence, each physical system corresponds to a unique Hilbert space.

\textbf{Postulate II:} An observable $O$ is a measurable property of a physical system. Mathematically, observables are represented by hermitian operators.

\textbf{Postulate III:} The eigenvalues $o_i$ of a hermitian operator $O = \sum_{i} o_i P_i$ represent the possible outcomes of measuring the observable $O$. Upon measuring $O$ and obtaining the result $o_i$, the quantum state $\ket{\psi}$ collapses to the state
\begin{align*}
	\frac{P_i\ket{\psi}}{\sqrt{p_i}},
\end{align*}
where $P_i$ is the projector onto the eigenspace associated with the eigenvalue $o_i$, and $p_i = \left\|P_i\ket{\psi}\right\|^2$ is the normalization factor.

\textbf{Postulate IV:} The probability $p_i$ of obtaining outcome $o_i$ when measuring the observable $O = \sum_{i} o_i P_i$ on a system described by the quantum state $\ket{\psi}$ is given by
\begin{align*}
	p_i = \text{Tr}\left(P_i \ket{\psi}\bra{\psi}\right) = \left\|P_i\ket{\psi}\right\|^2 = \braket{\psi|P_i|\psi}.
\end{align*}

Note that if $\ket{\psi'} = e^{i\theta}\ket{\psi}$ for some global phase $\theta \in \mathbb{R}$, the observable outcomes and probabilities remain unchanged. Therefore, quantum states differing only by a global phase are physically indistinguishable.

In relation to Postulate I, it is often convenient to represent quantum states using the \emph{density matrix} formalism, where the state is described by $\rho = \ket{\psi}\bra{\psi}$ instead of $\ket{\psi}$. In this formalism, the global phase ambiguity disappears, since if $\ket{\psi'} = e^{i\theta}\ket{\psi}$, then
\begin{align*}
	\rho' = \ket{\psi'}\bra{\psi'} = \ket{\psi}\bra{\psi} = \rho.
\end{align*}

Using this representation, Postulates III and IV are reformulated. The post-measurement state has a density matrix
\begin{align*}
	\frac{P_i \rho P_i}{p_i},
\end{align*}
and the probability of obtaining outcome $o_i$ is
\begin{align*}
	p_i = \text{Tr}(P_i \rho).
\end{align*}

Both formulations are equivalent, though one may be more convenient than the other depending on the context. Furthermore, the density matrix formalism naturally introduces the concept of \emph{mixed states}. Until now, we have implicitly dealt only with \emph{pure states}, i.e., states of the form $\rho = \ket{\psi}\bra{\psi}$.

Suppose now that, for reasons we ignore, the system is prepared in one of several pure states $\{\rho_i\}_{i=1}^n$ with probabilities $\{\eta_i\}_{i=1}^n$. Then the overall state of the system is described by the mixed state
\begin{align*}
	\rho = \sum_{i=1}^n \eta_i \rho_i.
\end{align*}

For the purposes of this project, we focus exclusively on pure states and do not address scenarios involving mixed states.

In relation to the observable postulate, we now define a \gls{povm} as a collection of measurement operators $\{O_i\}_{i=1}^n$ that satisfy,
\begin{itemize}
	\item Each $O_i$ is positive semi-definite, meaning all its eigenvalues $o_j^{(i)}$ satisfy $o_j^{(i)} \geq 0$.
	\item The operators satisfy the completeness relation,
\begin{align*}
	\sum_{i=1}^n O_i = \mathds{1},
\end{align*}
where $\mathds{1}$ denotes the identity operator on the Hilbert space $\mathcal{H}$.
\end{itemize}

Unlike projective measurements, the operators $O_i$ in a POVM are not necessarily orthogonal projectors. This allows the number of measurement outcomes to exceed the dimensionality of the Hilbert space, making POVMs particularly useful in tasks such as \gls{qse} and \gls{qsd}.

@@

\begin{align}\label{equationPerfectExclusionCondition}
	\lambda_{|G|} \leq \left( \sum_{i<|G|} \sqrt{\lambda_i} \right)^2.
\end{align}

@@

In Figure (\ref{FigureQSEMEGenericEigenValues3D}) we present a 3D plot for the comparison of the success probability for exclusion for ensembles of 3 states between generic and group-generated ensembles. The $x$ and $y$ axis representing de eigenvalues of the Gram matrix are analogous to previos figures. Nevertheless the $z$ axis shows the success probability for exclusion insted of the previos heatmaps. Here we can observe how the randomly generated states raise the probabilities over the group-generated values. The generation of unitary matrices that fill $U(n)$ ensures the Gram matrices we are generating comprehend the entire set of possible Gram matrices homogeniously.
\subsection{Intuition behind the results}

The intuition behind the exclusion task for group-generated ensembles being the worst case scenario for this task, thus yielding a lower bound for the probability of excluding a hypothesis state in the general case is as follows. Consider an ensemble $\{\ket{\psi_i}\}_{i=1}^n$ with a certain degree of symmetry, and a state $\ket{\psi_j}$ known to be a member of this ensemble. Our goal is to perform the exclusion task, i.e., to identify a state $\ket{\psi_k} \in \{\ket{\psi_i}\}_{i=1}^n$ such that we \emph{suspect} $\ket{\psi_k} \neq \ket{\psi_j}$\footnote{We emphasize the word \emph{suspect} because in many scenarios we are not certain about $\ket{\psi_k} \neq \ket{\psi_j}$, which corresponds to the \emph{non-perfect exclusion} case.}.

Due to the symmetric nature of the ensemble, the states are very similar to each other. This makes it especially difficult to identify which state is different from the target state. In contrast, if the ensemble includes a state $\ket{\tilde{\psi}}$ that is very different from the others, it becomes easier to determine whether $\ket{\tilde{\psi}}$ matches the target or not. Thus, when the symmetry is broken, exclusion can be performed more accurately.

To illustrate this more concretely, consider the following ensamble involving qubit states,
\[
\{\ket{\textcolor{red}{\uparrow}}, \ket{\textcolor{blue}{\uparrow}}, \ket{\textcolor{green}{\uparrow}}\}.
\]
These are represented on a Bloch circle\footnote{The Bloch circle is a projection of the Bloch sphere onto a plane.} in Figure~\ref{FigureIntuitionExample}.

\begin{figure}[H]
	\centering
	\begin{tikzpicture}[scale = 1]
	\draw[thick] (0,0) circle(1cm);
  	\draw[line width=1.5pt,->,>=stealth, red] (0,0) -- (0,1);
  	\draw[line width=1.5pt,->,>=stealth, blue] (0,0) -- (0.17365,-0.9848);
  	\draw[line width=1.5pt,->,>=stealth, green] (0,0) -- (-0.17365,-0.9848);
	\end{tikzpicture}
	\caption{Representation of the ensemble states as projections on the Bloch circle.}
	\label{FigureIntuitionExample}
\end{figure}

It is evident that $\ket{\textcolor{red}{\uparrow}}$ is noticeably different from the other two states. The red arrow points in the opposite direction to the blue and green arrows. Therefore, given an unknown state $\ket{\uparrow}$ from this ensemble, it is relatively easy to determine whether $\ket{\uparrow} = \ket{\textcolor{red}{\uparrow}}$ or $\ket{\uparrow} \neq \ket{\textcolor{red}{\uparrow}}$. If $\ket{\uparrow} = \ket{\textcolor{red}{\uparrow}}$, we can exclude either $\ket{\textcolor{blue}{\uparrow}}$ or $\ket{\textcolor{green}{\uparrow}}$; otherwise, we can exclude $\ket{\textcolor{red}{\uparrow}}$. In both cases, the exclusion is straightforward.

This is closely related to the task of quantum state discrimination (\gls{qsd}). If $\ket{\uparrow} = \ket{\textcolor{red}{\uparrow}}$, the exclusion and discrimination tasks are practically equivalent. However, if $\ket{\uparrow} \neq \ket{\textcolor{red}{\uparrow}}$, it becomes very difficult to determine whether $\ket{\uparrow} = \ket{\textcolor{blue}{\uparrow}}$ or $\ket{\uparrow} = \ket{\textcolor{green}{\uparrow}}$, as these states are almost indistinguishable. Thus, in this ensemble, the exclusion task may be more feasible than discrimination, showcasing its strength in certain scenarios.

%From an analytical perspective, consider the Gram matrix $\mathcal{G}$ for this ensemble. We have,
%\\begin{align*}
%\braket{\textcolor{blue}{\uparrow}|\textcolor{red}{\uparrow}} \approx \braket{\textcolor{green}{\uparrow}|\textcolor{red}{\uparrow}} = \epsilon, \quad \braket{\textcolor{blue}{\uparrow}|\textcolor{green}{\uparrow}} = 1 - \epsilon, \quad \text{with } |\epsilon|^2 \approx 0.
%\end{align*}
%The states are normalized, so,
%\begin{align*}
%\braket{\textcolor{red}{\uparrow}|\textcolor{red}{\uparrow}} = \braket{\textcolor{blue}{\uparrow}|\textcolor{blue}{\uparrow}} = \braket{\textcolor{green}{\uparrow}|\textcolor{green}{\uparrow}} = 1.
%\end{align*}
%Hence, the Gram matrix is approximately,
%\begin{align*}
%\mathcal{G} = 
%\begin{pmatrix}
%1 & \epsilon & \epsilon^* \\
%\epsilon^* & 1 & 1 - \epsilon \\
%\epsilon & 1 - \epsilon^* & 1
%\end{pmatrix}.
%\end{align*}
%Neglecting terms of order $\epsilon^2$, the eigenvalues of $\mathcal{G}$ are approximately,
%\begin{align*}
%\{2 + \Re(\epsilon), 1, -2\Re(\epsilon)\}.
%\end{align*}

Now contrast this with a symmetric ensemble, as shown in Figure~\ref{FigureIntuitionExample2}:

\begin{figure}[H]
	\centering
	\begin{tikzpicture}[scale = 1]
	\draw[thick] (0,0) circle(1cm);
  	\draw[line width=1.5pt,->,>=stealth, red] (0,0) -- (0,1);
  	\draw[line width=1.5pt,->,>=stealth, blue] (0,0) -- (0.866,-0.5);
  	\draw[line width=1.5pt,->,>=stealth, green] (0,0) -- (-0.866,-0.5);
	\end{tikzpicture}
	\caption{Symmetric ensemble represented on the Bloch circle.}
	\label{FigureIntuitionExample2}
\end{figure}

This configuration exhibits perfect symmetry. Define a unitary operator $\mathcal{U}$ such that:
\[
\mathcal{U} \ket{\textcolor{red}{\uparrow}} = \ket{\textcolor{blue}{\uparrow}},
\]
then by symmetry:
\[
\mathcal{U} \ket{\textcolor{blue}{\uparrow}} = \ket{\textcolor{green}{\uparrow}}, \quad \mathcal{U} \ket{\textcolor{green}{\uparrow}} = \ket{\textcolor{red}{\uparrow}}.
\]
Hence,
\[
\mathcal{U}^3 = \mathds{1}|_{\text{Bloch circle}},
\]
where $\mathds{1}|_{\text{Bloch circle}}$ denotes the identity operator restricted to this subspace. This implies that the ensemble is generated by a cyclic group of order 3, i.e., $\mathbb{Z}/3\mathbb{Z}$.

In such a group-generated ensemble, distinguishing any pair of states is equally difficult due to symmetry. Therefore, exclusion becomes more challenging, representing a worst-case scenario compared to the asymmetric case.

\section{Conclusions}

@@

I am grateful to Dr. Ramón Muñoz Tapia for allowing me to be one of his pupils and for presenting this project. I am also deeply grateful to PhD. Santiago Llorens Fernández for being the lighthouse that illuminates my path throughout this journey.
